In-silico molecular enrichment and clearance of the human intracranial space

### 2025-01-27 paper/supplementary_main.tex
Updates from Overleaf

We consider the system of coupled 3D-1D solute transport equations
given by~\eqref{eq:3d_pde} and~\eqref{eq:1d_pde} with the interface
conditions~\eqref{eq:3d_interface}, the bifurcation
conditions~\eqref{eq:conc_continuity}
and~\eqref{eq:flux_conservation}, and the boundary
conditions~\eqref{eq:3d_bcs} and~\eqref{eq:1d_bcs}. We discretize
these equations using an implicit finite difference scheme in time and
a discontinuous Galerkin (DG) finite element method with upwinding in space to accurately capture sharp boundary layers. We remark that the transport in the CSF spaces is highly convection dominated, with an average Péclet number of 402 and maximum of 9542 (assuming a characteristic length of $10\,$cm and taking the increased diffusivity due to dispersion into account).

@@

\begin{table}
\centering
\small
\begin{tabular}{ccccc}
\toprule
quantity of interest  & Watts et al\cite{watts2019measuring}  & Ringstad et al\cite{ringstad2018brain} & Eide and Ringstand \cite{eide2024functional} & ours (vasomotion model)\\
\midrule
Spinal transit time ($t_{st}$) &  1 - 3 hours & 20 min &  $13\pm6.3\,$ min & N/A \\
Time of peak amount (CSF) & 10 - 15 hours & 6–9 hours & N/A & 2 hours +$t_{st}$ \\
Time of peak amount (parenchyma) & 10 - 26 hours & 24 hours & N/A & 12 - 24 hours \\
Peak amount (CSF) &  N/A & N/A & N/A & 0.5 mmol \\
Peak amount (parenchyma) & N/A & N/A& N/A & 0.09 mmol \\
Injected tracer quantity &  1.0 mmol & 0.5 mmol & 0.5 mmol & 0.5 mmol (cranially) \\
Peak concentration (CSF) & 0.5 mmol/l &  N/A&  N/A & 1.4 mmol/l\\
Peak concentration (parenchyma) & 0.1 mmol/l & N/A & N/A & 0.4 mmol/l \\
FTA in MCA1 & N/A & N/A &  37.8 ± 47.0 min &  1:48 / 1:24 hours +$t_{st}$ \\
FTA in MCA2 & N/A & N/A & 53.1 ± 50.5 min & 2:12 / 1:48 hours +$t_{st}$ \\
FTA in ACA2 & N/A & N/A & 48.3 ± 46.1 min & 3:00 hours +$t_{st}$ \\
\bottomrule
\end{tabular}
\caption{Comparison of literature values and model predictions on tracer spreading patterns and timing}
\label{tab:spreading comparison}
\end{table}

>>>>

We consider the system of coupled 3D-1D solute transport equations
given by~\eqref{eq:3d_pde} and~\eqref{eq:1d_pde} with the interface
conditions~\eqref{eq:3d_interface}, the bifurcation
conditions~\eqref{eq:conc_continuity}
and~\eqref{eq:flux_conservation}, and the boundary
conditions~\eqref{eq:3d_bcs} and~\eqref{eq:1d_bcs}. We discretize
these equations using an implicit finite difference scheme in time and
a discontinuous Galerkin (DG) finite element method with upwinding in space to accurately capture sharp boundary layers. We remark that the transport in the CSF spaces is highly convection dominated, with an average Péclet number of 402 and maximum of 9542 (assuming a characteristic length of $10\,$cm, and accounting for the increased diffusivity due to dispersion).

@@

We compare our in-silico predictions with clinical data from human subjects, in particular the studies by Ringstad et al~\cite{ringstad2018brain}, Watts et al~\cite{watts2019measuring} and Eide and Ringstad~\cite{eide2024functional}. Finding an overall good agreement, we provide a detailed overview on key quantities of interest in \cref{tab:spreading_comparison}. We remark that the our simulation begins at the time of first tracer appearance at the craniocervical junction, while the clinical studies report time with respect to intrathecal injection time, leading to an offset, which is captured by the spinal transition time.

\begin{table}
\centering
\small
\begin{tabular}{lcccc}
\toprule
quantity of interest  & Watts et al\cite{watts2019measuring}  & Ringstad et al\cite{ringstad2018brain} & Eide and Ringstad \cite{eide2024functional} & high PVS flow model \\
\midrule
Spinal transit time ($t_{st}$) &  1 - 3 hours & 20 min &  $13\pm6.3\,$ min & N/A \\
Time of peak amount (CSF) & 10 - 15 hours & 6–9 hours & N/A & 2 hours +$t_{st}$ \\
Time of peak amount (parenchyma) & 10 - 26 hours & 24 hours & N/A & 12 - 24 hours \\
Peak amount (CSF) &  N/A & N/A & N/A & 0.5 mmol \\
Peak amount (parenchyma) & N/A & N/A& N/A & 0.09 mmol \\
Injected tracer quantity &  1.0 mmol & 0.5 mmol & 0.5 mmol & 0.5 mmol (cranially) \\
Peak concentration (CSF) & 0.5 mmol/l &  N/A&  N/A & 1.4 mmol/l\\
Peak concentration (parenchyma) & 0.1 mmol/l & N/A & N/A & 0.4 mmol/l \\
FTA in MCA1 & N/A & N/A &  37.8 ± 47.0 min &  1:48 / 1:24 hours +$t_{st}$ \\
FTA in MCA2 & N/A & N/A & 53.1 ± 50.5 min & 2:12 / 1:48 hours +$t_{st}$ \\
FTA in ACA2 & N/A & N/A & 48.3 ± 46.1 min & 3:00 hours +$t_{st}$ \\
\bottomrule
\end{tabular}
\caption{Comparison of clinical studies and model predictions on tracer spreading patterns and arrival times.}
\label{tab:spreading_comparison}
\end{table}
